---
title: "Pollution atmosphérique au dioxyde d'azote (no2) et à l'ozone (o3) ? dans le bassin grenoblois"
author: "FRANCES Tom, GAMBIEZ Jérôme, GOURDON Jérémie, MAHI Riad"
date: "Mars 2022"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Le contexte
The aim of the activity is to develop a methodology to answer a specific question on a given dataset. 

The dataset is the set of Firstname given in France on a large period of time. 
[https://www.insee.fr/fr/statistiques/2540004](https://www.insee.fr/fr/statistiques/fichier/2540004/dpt2020_txt.zip), we choose this dataset because it is sufficiently large, you can't do the analysis by hand, the structure is simple


You need to use the _tidyverse_ for this analysis. Unzip the file _dpt2020_txt.zip_ (to get the **dpt2020.csv**). Read in R with this code. Note that you might need to install the `readr` package with the appropriate command.

## Construction de la dataframe depuis les fichiers de données

Nous allons dans un premier temps récuperer les données depuis nos fichiers csv. 

```{r}
library(tidyverse)

data_crolles <- read_delim("Data/15048_Crolles.csv",delim=",", col_types = "ccTcdc")
data_crolles_f = data_crolles %>%
  filter(parameter == "no2");
data_crolles_f

data_voiron <- read_delim("Data/15018_Voiron.csv",delim=",", col_types = "ccTcdc")
data_voiron_f = data_voiron %>%
  filter(parameter == "no2")
data_voiron_f

data_rondeau <- read_delim("Data/15039_Rondeau.csv",delim=",", col_types = "ccTcdc")
data_rondeau_f = data_rondeau %>%
  filter(parameter == "no2")
data_rondeau_f

data_champ_sur_drac <- read_delim("Data/15013_Champ_sur_Drac.csv",delim=",", col_types = "ccTcdc")
data_champ_sur_drac_f = data_champ_sur_drac %>%
  filter(parameter == "no2")
data_champ_sur_drac_f

data_grenoble <- read_delim("Data/15046_Grenoble.csv",delim=",", col_types = "ccTcdc")
data_grenoble_f = data_grenoble %>%
  filter(parameter == "no2")
data_grenoble_f

```
Nombre d'observations pour chaque jeu de données (chaque ville)

```{r}
nb_observations_crolles = data_crolles_f %>% 
  summarize(nb_obs_crolles = n()) ;
nb_observations_crolles

nb_observations_voiron = data_voiron_f %>%
  summarize(nb_obs_voiron = n()) ;
nb_observations_voiron

nb_observations_rondeau = data_rondeau_f %>%
  summarize(nb_obs_rondeau = n()) ;
nb_observations_rondeau

nb_observations_champ_sur_drac = data_champ_sur_drac_f %>%
  summarize(nb_obs_champ_sur_drac = n()) ;
nb_observations_champ_sur_drac

nb_observations_grenoble = data_grenoble_f %>%
  summarize(nb_obs_grenoble = n()) ;
nb_observations_grenoble

```

Intervalle des dates

```{r}

range_date_crolles = data_crolles_f %>% 
  summarize(min_date_crolles = min(local), max_date_crolles = max(local)) ;
range_date_crolles

range_date_voiron = data_voiron_f %>%
  summarize(min_date_voiron = min(local), max_date_voiron = max(local)) ;
range_date_voiron

range_date_rondeau = data_rondeau_f %>%
  summarize(min_date_rondeau = min(local), max_date_rondeau = max(local)) ;
range_date_rondeau

range_date_champ_sur_drac = data_champ_sur_drac_f %>%
  summarize(min_date_champ_sur_drac = min(local), max_date_champ_sur_drac = max(local)) ;
range_date_champ_sur_drac

range_date_grenoble = data_grenoble_f %>%
  summarize(min_date_grenoble = min(local), max_date_grenoble = max(local)) ;
range_date_grenoble


```

Moyenne des quantitées de no2 par jour


```{r}
FirstNames %>% 
  filter(preusuel == "TOM") %>% 
  group_by(preusuel, annais) %>% 
  summarise(total = sum(nombre)) %>%
  ggplot(aes(x=annais, y=total)) + geom_point() + ylim(0,NA) + theme_bw();

```
```{r}
nb_NA_observations = FirstNames %>% filter(is.na(annais) | is.na(dpt)) %>% count();
nb_NA_observations
nb_observations = FirstNames %>% count();
nb_observations
nb_NA_observations * 100 / nb_observations;
```
Number of _PRENOMS_RARES
```{r}
n_rare = FirstNames %>%
  filter(preusuel == "_PRENOMS_RARES") %>%
  group_by(sexe) %>%
  summarise(n_rare = n())
n_rare
```



GGPLOT fonctionnement
```{r}
library(ggplot2)
ggplot()
pressure
ggplot(data = pressure) + 
  aes(x=temperature) + 
  aes(y=pressure) + 
  geom_point() + 
  geom_line() +
  labs(x="Temperature (deg C)") +
  labs(y = "Pression") +
  labs(title = "Titre")
```

